---
title: "Assignment-01"
author: "Yifan Meng"
date: 2025/2/16
format: 
  html:
    embed-resources: true
    toc: true
    number-sections: true
    highlight-style: github
editor: visual
---

## Question 1 \[12pts\]

A performance test of manual dexterity was administered to a group of subjects. The frequency distribution is reported as follows.

| Score     | 50  | 60  | 70  | 80  | 90  |
|-----------|-----|-----|-----|-----|-----|
| Frequency | 1   | 2   | 3   | 3   | 1   |

### Part 1 \[3pts\]

#### Question

Compute the mean, variance, and standard deviation of the score distribution.

[*Hint: You can use mean(), var(), and sd() functions in R.*]{style="color: grey;"}

#### Answer

```{r}
#| label: q1-stats
#| message: false

# Create the data
scores <- rep(c(50, 60, 70, 80, 90), times=c(1, 2, 3, 3, 1))

# Calculate statistics
mean(scores)
var(scores)
sd(scores)
sd(scores)/sqrt(length(scores))

## Values of the answer below is calculated by function round(), e.g.: round(sd(scores),2)

```

::: {.callout-note title="Report"}
-   The mean of this score distribution is `r round(mean(scores),2)`

-   The variance of the score distribution is `r round(var(scores),2)`

-   The standard deviation of the score distribution is `r round(sd(scores),2)`
:::

### Part 2 \[9pts\]

#### Question

The following table contains raw scores, z-scores, and T-scores for Joan and Peter from the group. Complete the table by finding out the unknown values (a), (b), and (c).

| Name  | Raw score | z-score | T-score |
|-------|-----------|---------|---------|
| Joan  | 80.00     | 0.75    | 57.50   |
| Peter | 49.45     | -1.80   | 32.00   |

#### Answer

::: {.callout-tip title="Solution Process"}
According to **Part 1**, $\mu$ estimated by the distribution is `r round(mean(scores),2)` and $\sigma$ is `r round(sd(scores),2)`.
:::

##### (a) Joan's z-score:

$$
\begin{aligned}
z &= \frac{X - \mu}{\sigma} \\
z &= \frac{80 - 71}{11.97} = \frac{9}{11.97} \approx 0.75
\end{aligned}
$$

##### (b): Peter's raw score

$$
X = z \cdot \sigma + \mu
$$

Substituting the values for Peter's z-score:

$$
\begin{aligned}
X &= (-1.80) \cdot 11.97 + 71 \\
X &= -21.53 + 71 = 49.47
\end{aligned}
$$

##### (c): Peter's T-score:

The formula of T score is : $$
T = 10\cdot z+50
$$ Calculate the Joan's T score by z-score $$
T = 10\cdot(-1.80)+50=32
$$

## Question 2 \[11pts\]

James is interested in finding the relationship between Introversion and Social Dominance. He collected the following data and calculated the descriptive statistics of the data.

| Examinee         | 1   | 2   | 3   | 4   | 5   | 6   | 7   | 8   | Mean | SD   |
|------------------|-----|-----|-----|-----|-----|-----|-----|-----|------|------|
| Introversion     | 0   | 1   | 1   | 2   | 2   | 3   | 4   | 5   | 2.25 | 1.67 |
| Social Dominance | 4   | 5   | 4   | 2   | 3   | 3   | 2   | 1   | 3.00 | 1.31 |

### Part 1 \[3pts\]

#### Question

Obtain a scatter plot of the two variables by using plot() function in R. Please put Introversion on the x-axis and Social Dominance on the y-axis.

[*Hint: You can construct the dataset in .csv or .txt format and import the data OR you can also create data objects by running the codes below:*]{style="color: grey;"}

#### Answer

```{r}
#| label: fig-scatterplot
#| fig-align: center
#| fig-width: 7
#| fig-height: 7
#| echo: true
#| message: true
#| warning: false
Introv <- c(0, 1, 1, 2, 2, 3, 4, 5) 
Socialdom <- c(4, 5, 4, 2, 3, 3, 2, 1)
#df <- data.frame(Introv, Socialdom)

plot(Introv,Socialdom,main = "Scatter Plot of Introversion and Social Dominance",
     xlab = "Introversion", ylab = "Social Dominance")

```

### Part 2 \[6pts\]

#### Question

Based on the scatter plot you obtained in 2-1), how would you describe the: i) trend ii) strength iii) direction of the association between the two variables? Select your response from the following options for each element:

a)  Positive b) Negative
b)  Linear d) Non-linear
c)  Strong f) Weak

#### Answer

iii) direction: b) Negative
iv) strength: e) Strong
v)  trend: c) Linear

::: {.callout-note title="Details"}
-   **Direction (Negative)**: As Introversion increases, Social Dominance tends to decrease.

-   **Strength (Strong)**: The points are closely clustered around a downward trend, indicating a strong relationship (also can be supported by the correlation coefficient below).

-   **Trend (Linear)**: The pattern follows a straight line rather than a curve.
:::

### Part 3 \[2pts\]

#### Question

Compute the Pearson correlation coefficient between Introversion and Social Dominance using cor() function in R.

#### Answer

```{r}
#| label: q2-correlation
#| message: false
cor(Introv,Socialdom)
```

The Pearson correlation between Introversion and Social Dominance is `r round(cor(Introv,Socialdom),2)`

## Question 3 \[8pts\]

Determine the scale of measurement for each of the following variables.

1)  \[2pts\] Ranking of baseball teams (determined by their won-lost records at the end of the season): **Ordinal Scale**

2)  \[2pts\] Social Security numbers: **Nominal Scale**

3)  \[2pts\] Air distance between New York City and other cities in the United States: **Interval Scale**

4)  \[2pts\] pH value: **Ratio Scale**

## Question 4 \[9pts\]

The scores of a reference population on the Wechsler Intelligence Scale for Children (WISC) are normally distributed with μ = 100 and σ = 15.

#### Question 1 \[3pts\]

Approximately what percentage of the population would score below 100 on the test? **50%**

#### Answer 1

::: {.callout-note title="Solution"}
The WISC scores follow a normal distribution:

$$X \sim \mathcal{N}(100, 15^2)
$$

Here we want to estimate $$
    P(X < 100)
$$ Because $\mu$=100, so z-score of 100 is 0.

From the standard normal table, the probability of a z-score being less than 0 is 50%
:::

Or directly calculate using R

```{r}
    pnorm(100, mean = 100, sd = 15)
```

#### Question 2 \[3pts\]

Approximately what percentage of the population would score above 130 on the test?

#### Answer 2

The z score of 130 in this test is 2, so there will be approximately 2.5% of the population would score above 130 on the test.

::: {.callout-note title="Solution"}
Here we want to estimate: $$
    P(X > 130)
$$

Then, calculate the z-score for 130:

$$
    \begin{aligned}
    z &= \frac{130 - 100}{15} \\[1em]
    z &= \frac{30}{15} = 2
    \end{aligned}
$$ Based on the standard norm table:

$$
    P(Z > 2) = 1 - P(Z < 2) \approx 1-0.975 \approx 0.025
$$

For z = 2:

-   Area above z = 2 ≈ 2.5%
-   Therefore, approximately **2.5%** of the population would score above 130
:::

Or directly calculate using R

```{r}
    1-pnorm(130,mean=100,sd=15)
    #or
    pnorm(130,mean=100,sd=15,lower.tail = FALSE)
```

#### Question 3 \[3pts\]

Approximately what percentage of the population would score between 100 and 130 on the test? **47.5%**

#### Answer 3

::: {.callout-note title="Solution"}
Here we want to calculate: $$
    P(100<X<130)
$$ From 2 subquestions above,

$$
\begin{aligned}
    P(X < 130) &\approx 0.975\\
    P(X > 100)&=0.5
\end{aligned}
$$ So:

$$
P(100<X<130)=P(X < 130)-P(X > 100)=0.975-0.5=47.5%
$$
:::

In R:

```{r}
    pnorm(130, mean = 100, sd = 15) - pnorm(100, mean = 100, sd = 15)
```

```{r}
#| label: fig-density
#| fig-align: center
#| fig-width: 7
#| fig-height: 5
#| echo: false
#| message: true
#| warning: false
x <- seq(40, 160, length=300)
y <- dnorm(x, mean=100, sd=15)

plot(x, y, type="l", lwd=2, col="blue", 
     xlab="", ylab="", 
     xaxt="n",
     yaxt="n",
     main="")

sigma_points <- c(-3,-2,-1,0,1,2,3) * 15 + 100
axis(1, at=sigma_points,
     labels=c("-3σ","-2σ","-1σ","μ","+1σ","+2σ","+3σ"))

abline(v=sigma_points, col="gray", lty=2)

# fill
# 68%  (μ±1σ)
polygon(c(x[x>=85 & x<=115], 115, 85), 
        c(y[x>=85 & x<=115], 0, 0), 
        col=rgb(0,0,1,0.2))

# 95%  (μ±2σ)
polygon(c(x[x>=70 & x<=130], 130, 70), 
        c(y[x>=70 & x<=130], 0, 0), 
        col=rgb(0,0,1,0.1))

# 99.7%  (μ±3σ)
polygon(c(x[x>=55 & x<=145], 145, 55), 
        c(y[x>=55 & x<=145], 0, 0), 
        col=rgb(0,0,1,0.05))

# percentile
text(100, 0.018, "68%", cex=1.2)
text(100, 0.022, "95%", cex=1.2)
text(100, 0.026, "99.7%", cex=1.2)

lines(x, y, lwd=2, col="blue")

```
